---
title: Example I
description: ''
date: 2025-06-04
weight: 5
---

```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

```{r, echo=FALSE, include=FALSE}
source("R/site_globals.R")
source("scripts/geomUtils.R")
source("scripts/config.R")
library(leaflet)
library(rmarkdown)
library(dplyr)
library(ggplot2)
```

This map shows the accumulated search effort for Primula pauciflora over both historical and potential habitat:

```{r}
prim_agm <- read.csv("Analysis_outputs/Intermediate/Primula pauciflora_accepted_grouped_merged.csv")
prim_sf = assign_cell_geometry_sf(prim_agm, galgrid)

pal <- colorNumeric(palette = "viridis", domain = range(c(0, prim_sf$search_effort), na.rm = TRUE))
m <- leaflet(data = prim_sf) %>%
  addProviderTiles("Esri.WorldImagery") %>%
  # Add the grid layer to the map
  addPolygons(fillColor = ~pal(search_effort), fillOpacity = 0.8, 
              color = "#BDBDC3", weight = 1) %>%
  # Add a legend
  addLegend(pal = pal, values = c(0, max(prim_sf$search_effort, na.rm = TRUE)),
            opacity = 0.8, title = "Accumulated Search Effort in ks")

# Print the map
m

```

See the treatment of [historical habitat](../../extirpation_historical/example-i) for explanation of
the data format underlying this plot.


This produces posterior regional statistics for extirpation across all habitat as follows:

```{r, message=FALSE, warning=FALSE, echo=FALSE}
target_stats <- read.csv("Analysis_outputs/Intermediate/Primula pauciflora_stats.csv")
target_stats_all <- target_stats %>% dplyr::filter(Population == "all")
paged_table(target_stats_all)
```

These show the computed parameters of the posterior beta distribution for extirpation expressed in two different forms - the standard (alpha, beta) representation and (mu, var) as parameters for the central estimate for sighting probability and its dispersion. Confidence bands are placed for this at [`r target_stats_all$Low`, `r target_stats_all$High`].

Here is this posterior distribution graphed out - whilst the most probable extirpation value is 100% because of the folded nature of the beta distribution, this is a more dispersed distribution than that for historical habitat and the central estimate of extirpation probability is `r target_stats_all$Central`.

```{r, message=FALSE, warning=FALSE, echo=FALSE}
plot_site_beta(target_stats_all, "Primula pauciflora")
```

