---
title: Target selection criteria
description: Honing focus on biodiversity change
tags: [test, docs]
weight: 4
---

```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

```{r, echo=FALSE, include=FALSE}
source("R/site_globals.R")
source("scripts/geomUtils.R")
source("scripts/config.R")
library(rmarkdown)
library(dplyr)
library(ggplot2)
```


## Open-ended assessments of potential change

In order to proceed beyond an initial baseline analysis to assess possible changes in a system, we can set
criteria to focus our attention (and limited resources) on high risk species that are most likely to be vulnerable.

Depending on the criteria we set, there may emerge many potential target species or populations of concern.
On the first pass, however, it is notable that this decision-making process is blind to any conservation priorities 
we might have (*e.g.*, species assessed as vulnerable at a provincial or national scale), which can bias biodiversity 
change assessments.

{{% pageinfo %}}
"Conservation programs often prioritize species-at-risk, even while many common species may undergo the highest relative losses (Jansen *et al.*, 2019)."
{{% /pageinfo %}}


## Criteria to consider

### Provenance

First, presumably we are mostly concerned with change in plant communities of native origin:

```{r, message=FALSE, warning=FALSE, results='hide'}
native.at.large.2020 <- at.large.2020 %>% filter(Origin == 'native')
nrow(native.at.large.2020) # 110 native plant species at large
```

### Historical evidence

Biological specimens are the strongest evidence we have for biodiversity change assessments.

```
# Filter voucher specimens
# 
```

### Taxonomically difficult or cryptic species

```
# Filter out graminoids
# 
```


### Coordinate uncertainty

Geographic coordinates represent one of the most crucial pieces of information for biodiversity change assessments.

* Often we are confronted with limited locality information, which imposes constraints on representation of historical habitat
* Sometimes, the information is good enough to map species occurrences to precise locations (*e.g.*, 30x30m2 grid cell)
* Sometimes, the information is good enough to specify a discrete location or habitat patch
* Sometimes, coordinates are generalized to the geometric centre (centroid) of a given area, with a wide radius of coordinate uncertainty
* Sometimes, we don't have any coordinates at all

It is exceedingly difficult if not impossible to assess the status of species or populations that lack specific locality information. At the very least, we will want to ensure our targets have coordinates:

```
# Filter out records lacking coordinates
```

(Depending on the approach, one might want to be more strict and filter records below a certain threshold of coordinate Uncertainty)

